---
title: "Vignette for metacell co‐accessibility analysis with RWireX"
author: "Isabelle Seufert"
output: html_document
---
# Load libraries
library(ArchR)
library(plotgardener)
library(SpectralTAD)

library(RWireX)

# Load exemplary data


# Generate tile accessibility matrix
For now, we generate the tile accessibility matrix only for chromosome 6 to reduce run time and required resources. We selected 10 kb for tile sizes, but this can be adapted considering data quality and resolution requirements.

```{r}
### Exclude all chromosomes, except chromosome 6
excludeChr <- getSeqnames(proj)
excludeChr <- c(excludeChr, "chrM", "chrY")
excludeChr <- excludeChr[excludeChr != "chr6"]

### Generate tile accessibility matrix without binarizing the counts
proj <- addTileMatrix(proj, tileSize = 10000, binarize = FALSE, 
                      excludeChr = excludeChr, force = TRUE)
```

# Compute metacell co-accessibility
XXX Next, we compute metacell co-accessibility across all samples to , as the samples represent cell from one TNFa treatment time point selected for a homogeneous population of cells in G1 cell cycle state. If you use your own data, you might want to consider selecting homogeneous cell clusters or cell types for separate analyses.

```{r}
### Define number of cells per replicate
numCellsPerAggregate <- 10

### Define number of aggregates
totalCells <- nrow(proj@cellColData)
numAggregates <- 0.9 * totalCells / numCellsPerAggregate

### Run co-accessibility analysis
mcCoacc <- RWireX::getCoAccessibility(proj, 
                                      useMatrix = "TileMatrix", 
                                      binaryMatrix = FALSE,
                                      maxDist = 2e+06, 
                                      AggregationMethod = "unique",
                                      numCellsPerAggregate = numCellsPerAggregate,
                                      numAggregates = numAggregates, 
                                      log2Norm = FALSE, 
                                      returnLoops = TRUE)
}
```

# Call contiguous domains of enriched co-accessibility (DCs)
XXX

```{r}
DCs <- RWireX::getCoAccessibleDomains(mcCoacc)
}
```

Have a look at the number of DCs.

```{r}
length(DCs)
```

We can have a look at the distribution of background co-accessibility per sample and select appropriate cutoffs. We recommend using the 99th percentile of background co-accessibility as lower cutoff to filter for true-positive co-accessible links.
```{r}
p <- ggplot()

for (sample in samples){
    df <- rbind(bgCoacc_list[[sample]]$BackgroundCoAccessibility$featShuffle,
                bgCoacc_list[[sample]]$BackgroundCoAccessibility$cellShuffle)
    df$sample <- sample
    
    p <- p + geom_histogram(data = as.data.frame(df),
             aes(x = correlation, fill = sample), alpha = 0.5)
}

p
```

# Visualize DCs by co-accessibility map
We visualize DCs by heatmap. The heatmap color reflects the correlation coefficient. DCs called from metacell co-accessibility are annotated. Here, we selected an exemplary region on chromosome 6 from our publication.

```{r}
DC_example <- as("chr6:136573673-137633313", "GRanges")

plotCoAccessibilityMap(CoAccessibility = mcCoacc,
                       region = DC_example,
                       annotation = DCs,
                       onlyPos = FALSE, rescale = FALSE)
}
```