---
title: "Vignette for metacell co‐accessibility analysis with RWireX"
author: "Isabelle Seufert"
output: html_document
---

### Load libraries
```{r}
library(ArchR)
library(plotgardener)
library(SpectralTAD)

library(RWireX)
```

### Set ArchR parameters
```{r}
addArchRThreads(threads = 10)
addArchRGenome("hg38")
```

### Download exemplary data
```{r}
data_paths <- downloadVignetteData()
```

### Import test data
```{r}
proj <- loadArchRProject(data_paths$proj)
PeakSet <- readRDS(data_paths$peaks)
```

### Get samples
```{r}
samples <- unique(proj$Sample)
```

# Generate tile accessibility matrix
For now, we generate the tile accessibility matrix only for chromosome 6 to reduce run time and required resources. We selected 10 kb for tile sizes, but this can be adapted considering data quality and resolution requirements.

```{r}
### Exclude all chromosomes, except chromosome 6
excludeChr <- getSeqnames(proj)
excludeChr <- c(excludeChr, "chrM", "chrY")
excludeChr <- excludeChr[excludeChr != "chr6"]

### Generate tile accessibility matrix without binarizing the counts
proj <- addTileMatrix(proj, tileSize = 10000, binarize = FALSE, 
                      excludeChr = excludeChr, force = TRUE)
```

# Compute metacell co-accessibility
Next, we compute metacell co-accessibility across all samples. If you use your own data, you might want to consider selecting multiple cell clusters or cell types for combined analysis.

```{r}
### Define number of cells per metacell
numCellsPerMetacell <- 10

### Define number of metacells
totalCells <- nrow(proj@cellColData)
numMetacells <- 0.9 * totalCells / numCellsPerAggregate

### Run co-accessibility analysis
mcCoacc <- RWireX::getCoAccessibility(proj, 
                                      useMatrix = "TileMatrix", 
                                      binaryMatrix = FALSE,
                                      maxDist = 2e+06, 
                                      AggregationMethod = "unique",
                                      numCellsPerAggregate = numCellsPerMetacell,
                                      numAggregates = numMetacells, 
                                      log2Norm = FALSE, 
                                      returnLoops = TRUE)
```

# Call contiguous domains of enriched co-accessibility (DCs)
From the metacell co-accessibility, we can identify domains of contiguously enriched co-accessibility. This will take a few minutes to run.

```{r}
DCs <- RWireX::getCoAccessibleDomains(mcCoacc)
```

Have a look at the called DCs.

```{r}
DCs
```

# Visualize DCs by co-accessibility map
We can visualize DCs by heatmap. The heatmap color reflects the correlation coefficient from metacell co-accessibility. DCs called from metacell co-accessibility are annotated. Here, we selected an exemplary region on chromosome 6 from our preprint.

```{r}
DC_example <- as("chr6:137100000-138100000", "GRanges")

plotCoAccessibilityMap(CoAccessibility = mcCoacc,
                       region = DC_example,
                       annotation = list(DCs = DCs),
                       onlyPos = FALSE, rescale = FALSE)
}
```