---
title: "Vignette for single cell co‐accessibility analysis with RWireX"
author: "Isabelle Seufert"
output: html_document
---

### Load libraries
```{r}
library(ArchR)
library(plotgardener)
library(SpectralTAD)

library(RWireX)
```

### Set ArchR parameters
```{r}
addArchRThreads(threads = 10)
addArchRGenome("hg38")
```

### Download exemplary data
```{r}
data_paths <- downloadVignetteData()
```

### Import test data
```{r}
proj <- loadArchRProject(data_paths$proj)
PeakSet <- readRDS(data_paths$peaks)
```

### Get samples
```{r}
samples <- unique(proj$Sample)
```

# Generate peak accessibility matrix
First, we generate the peak accessibility matrix. For now, we subset our peak set to chromosome 8 to reduce run time and required resources.

```{r}
### Only take peaks on chromosome 8
PeakSet <- PeakSet[seqnames(PeakSet) == "chr8"]

### Add peaks to ArchR project
proj <- addPeakSet(proj, peakSet = PeakSet)

### Generate peak accessibility matrix without binarizing the counts
proj <- addPeakMatrix(proj, binarize = FALSE)
```

# Compute single cell co-accessibility
Next, we compute single cell co-accessibility for each sample separately, as each sample consists of cells from one TNFa treatment time point. For each sample, we previously selected for a homogeneous population of cells in G1 cell cycle state. If you use your own data, you might want to consider selecting homogeneous cell clusters or cell types for separate analyses.

```{r}
scCoacc_list <- list()

for (sample in samples){
    ### Get all cells in sample
    cells <- rownames(proj@cellColData)[proj$Sample == sample]

    ### Compute co-accessibility from selected cells
    scCoacc_list[[sample]] <- RWireX::getCoAccessibility(proj, 
                                                         useMatrix = "PeakMatrix",
                                                         cellsToUse = cells,
                                                         maxDist = 1e+06, 
                                                         AggregationMethod = "single_cell_resolution",
                                                         log2Norm = FALSE,
                                                         returnLoops = TRUE)
}
```

# Compute background co-accessibility
To assess which of these co-accessible links are true-positive chromatin interactions, we compute the background co-accessibility for each sample separately. 

```{r}
bgCoacc_list <- list()

for (sample in samples){
    ### Get all cells in sample
    cells <- rownames(proj@cellColData)[proj$Sample == sample]

    ### Compute co-accessibility from selected cells
    bgCoacc_list[[sample]] <- RWireX::getBackgroundCoAccessibility(proj, 
                                                                   useMatrix = "PeakMatrix",
                                                                   cellsToUse = cells,
                                                                   maxDist = 1e+06, 
                                                                   AggregationMethod = "single_cell_resolution",
                                                                   log2Norm = FALSE)
}
```

# Filter autonomous links of enriched co-accessibility (ACs)
We can have a look at the distribution of background co-accessibility per sample and select appropriate cutoffs. We recommend using the 99th percentile of background co-accessibility as lower cutoff to filter for true-positive co-accessible links.

```{r}
p <- ggplot() + theme_minimal()

for (sample in samples){
    df <- rbind(bgCoacc_list[[sample]]$BackgroundCoAccessibility$featShuffle,
                bgCoacc_list[[sample]]$BackgroundCoAccessibility$cellShuffle)
    df$sample <- sample
    
    p <- p + geom_density(data = as.data.frame(df),
             aes(x = correlation, color = sample)) +
             geom_vline(xintercept = bgCoacc_list[[sample]]$BackgroundCutoff)
}

p
```

Have a look at the 99th percentile cutoffs from background co-accessibility for each sample.
```{r}
lapply(bgCoacc_list, function(x){round(x$BackgroundCutoff, 4)})
```

Now, we can filter for high-confidence co-accessible links. We select high-confidence co-accessible links using a lower cutoff for correlation coefficients from the background distribution, a lower cutoff for percent accessible cells (frequency of link occurence in single cells) and only selecting links with positively correlated peak accessibility.
```{r}
AC_list <- list()

for (sample in samples){
    AC_list[[sample]] <- RWireX::filterCoAccessibility(coAccessibilityLoops = scCoacc_list[[sample]], 
                                                       corCutOff = bgCoacc_list[[sample]]$BackgroundCutoff, 
                                                       onlyPos = TRUE,
                                                       perAccess = 5)
}
```

Have a look at the number of ACs per sample.
```{r}
lapply(AC_list, function(x){length(x$CoAccessibility)})
```

# Visualize ACs by loops
We visualize ACs by loops. The color of loops reflects the correlation coefficient, while the height of loops reflects the frequency of the AC occurence in single cells. Here, we selected an exemplary region on chromosome 8 from our preprint.

```{r}
AC_example <- as("chr8:102573784-102880000 ", "GRanges")

p <- RWireX::plotBrowserTrack(proj, 
                              region = AC_example,
                              groupBy = "Sample", 
                              normMethod = "nFrags", 
                              loops = lapply(AC_list, function(x){x$CoAccessibility}))
grid.newpage(); grid.draw(p)
```

# Compute AC activities per cell

```{r}
AC_activities <- getLinkActivityMatrix(proj,
                                       AC_list$HUVEC_TNFa_30min_Rep1, 
                                       regionSubset = AC_example)
```

We can visualize AC activities by heatmap and find that some co-accessible links are active in almost all cells, while others appear to be less frequent. We find that some of these rare co-accessible links are commonly active together.
```{r}
### Get dataframe of AC activities
df <- reshape2::melt(as.matrix(AC_activities))

### Cluster cells by AC activity
cluster_cells <- hclust( dist(t(AC_activities), method = "euclidean"), method = "ward.D2" )$order

### Reorder cells in activitydataframe
df$Var2 <- factor(df$Var2, levels = colnames(AC_activities)[cluster_cells])
```

```{r}
### Plot heatmap of AC activities
p <- ggplot(df, aes(x = Var2, y = Var1, fill = value)) + 
          geom_tile() + xlab("Cells")
p
```